High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2020-06-06 paper/main.tex
updates to paper and to pca level analysis

\begin{align}
Y_{n+1,t} = vec(\mathbf{\Sigma}_{n,t})\cdot vec(\mathbf{\Sigma}_{n,t})^\top\label{eqn:veccorr}
  \end{align}

 Observations were then drawn at each timepoint from a  ${ Y_{n,t}}$-mean multivariate Gaussian distribution with a
covariance matrix defined as ${ Y_{n+1,t}}$.

@@

\subsection*{Recovering known higher-order dynamic correlations from synthetic data}
We generated synthetic datasets that differed in how the underlying
correlations changed over time.  For each dataset, we applied
Equation~\ref{eqn:timecorr} with a variety of kernel shapes and
widths.  We assessed how well the true underlying correlations at each
timepoint matched the recovered correlations
(Fig.~\ref{fig:synthetic}).  For every kernel and dataset we tested,
our approach recovered the correlation dynamics we embedded into the
data.  However, the quality of these recoveries varied across
different synthetic datasets in a kernel-dependent way.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first and second order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    correlation matrix and the simulated correlation matrix at each
    timepoint and for each synthetic dataset.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see \textit{Synthetic data}.  \textbf{a. Constant correlations.}
    These first and second order dynamic correlation were
    simulated from datasets have a stable (unchanging) underlying correlation
    matrix.  \textbf{b. Random correlations.}  These first and second order dynamic correlation were
    simulated from datasets are
    generated using a new independently drawn correlation matrix at
    each new timepoint.  \textbf{c. Ramping correlations.}  These first and second order dynamic correlation were
    simulated from
    datasets are generated by smoothly varying the underlying
    correlations between the randomly drawn correlation matrices at
    the first and last timepoints and displays the
    correlations between the recovered dynamic correlations and the
  simualted correlation first and second order correlation matrices.   \textbf{d. Event-based correlations.}  These first and second order dynamic correlation were
    simulated from datasets are
    each generated using five randomly drawn correlation matrices that
    each remain stable for a fifth of the total timecourse.  }
  \label{fig:higherr_order_sims}
\end{figure}

>>>>

\begin{align}
Y_{n=1,t} = vec(\mathbf{\Sigma}_{t})\cdot
  vec(\mathbf{\Sigma}_{t})^\top \\
  Y_{n+1,t} = vec(Y_{n,t}) \cdot vec(Y_{n,t})^\top\label{eqn:veccorr}
  \end{align}

  We repeated this processes for the number of desired orders (N) to
  generate a timeseries of $\mathbf{N^{th}}$ correlation matrices.  Then, observations were drawn at each timepoint from a  ${ Y_{N-1,t}}$-mean multivariate Gaussian distribution with a
 covariance matrix defined as ${ Y_{N+1,t}}$ and the resultant vector
 was reshaped to the matrix the size of ${ Y_{N-1,t}}$:
 \begin{align}
   { \hat{Y}_{N-1,t}} \sim vec(\mathcal{N}({ Y_{n,t}}, { Y_{N,t}}))^{-1}
   \end{align}
 This process was then repeated for all decreasing orders, in which
 observations werre drawn at each timepoint from a  ${ Y_{n,t}}$-mean
 multivariate Gaussian distribution, but the covariance matrix is
 defined at the previously estimated reshaped matrix $\hat{Y}_{n,t}$:
 \begin{align}
   { \hat{Y}_{n,t}} \sim vec(\mathcal{N}({ Y_{n,t}}, {\hat{ Y}_{n+1,t}}))^{-1}
 \end{align}
We repeat this process until $n = 1$, then we drew observations at each timepoint
from a zero-mean multivariate Gaussian distribution with a
covariance matrix defined for each timepoint as described below:

\begin{align}
 s_t \sim \mathcal{N}(\mathbf{0},  \hat{Y}_{n=1,t})
\end{align}

For our specific analyses the resulting synthetic timeseries comprised
50 features and 300 timepoints. We then used this synthetic data to test how well Equation~\ref{eqn:timecorr}
could recover the simulated first and second order correlations.  To mitigate temporal blur in
estimating each timeseries of dynamic correlations, we used a Dirac $\delta$ function
kernel (which places all of its mass over a single timepoint)
Fig.~\ref{fig:kernels}b) to compute each lower-order correlation and then used a
Laplace kernel with a width of 20 to compute  $\mathbf{ X_{n,t}}$, applying
temporal blur at just the last step.  Then, for each timepoint, we
correlated the simulated correlation matix, ${ Y_{n,t}}$ with the
Equation~\ref{eqn:timecorr} estimated correrlation matric, $\mathbf{ X_{n,t}}$.

@@

\subsection*{Recovering known higher-order dynamic correlations from synthetic data}
We generated a synthetic timeseries from simulated first and second order dynamic correlation
matrices that differed in how the underlying
correlations changed over time.  For each dataset, we applied
Equation~\ref{eqn:timecorr} with a Laplace kernel with a width of 20.
We assessed how well the recovered correlations matched simulated
first and second order correlations. 
(Fig.~\ref{fig:higher_order_sims}).  For every dataset we tested,
our approach recovered the correlation dynamics we embedded into the
data.  However, the quality of these recoveries varied across
different synthetic datasets.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first and second order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    correlation matrix and the simulated correlation matrix at each
    timepoint and for each synthetic dataset.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see \textit{Synthetic data}.  \textbf{a. Constant correlations.}
    These first and second order dynamic correlation were
    simulated from datasets have a stable (unchanging) underlying correlation
    matrix.  \textbf{b. Random correlations.}  These first and second order dynamic correlation were
    simulated from datasets are
    generated using a new independently drawn correlation matrix at
    each new timepoint.  \textbf{c. Ramping correlations.}  These first and second order dynamic correlation were
    simulated from
    datasets are generated by smoothly varying the underlying
    correlations between the randomly drawn correlation matrices at
    the first and last timepoints and displays the
    correlations between the recovered dynamic correlations and the
  simualted correlation first and second order correlation matrices.   \textbf{d. Event-based correlations.}  These first and second order dynamic correlation were
    simulated from datasets are
    each generated using five randomly drawn correlation matrices that
    each remain stable for a fifth of the total timecourse.  }
  \label{fig:higher_order_sims}
\end{figure}
